---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# exactamente

<!-- badges: start -->
<!-- badges: end -->

`exactamente` is a versatile R package that offers a collection of tools to assist researchers and data analysts in exploring different bootstrap methods on small sample size data. Bootstrap methods are widely used for estimating the sampling distribution of an estimator by resampling with replacement from the original sample. 

This package is focused particularly on the exact case bootstrap and the exact bootstrap as described in [Jekel & Romero (2020)](https://doi.org/10.2172/1605343). Both methods are advantageous for bootstrapping small data sets where standard methods might be inadequate.

For a given sample size N, the exact case bootstrap generates all (2N - 1) choose N resamples. For instance, for the sample [1,2,7] as used in the Jekel & Romero paper, we would see 10 unique resamples:

[1, 1, 1] [1, 1, 2] [1, 1, 7] [1, 2, 2] [1, 2, 7]
[1, 7, 7] [2, 2, 2] [2, 2, 7] [2, 7, 7] [7, 7, 7].

In contrast, the exact bootstrap generates N^N resamples, thereby including permutations such as [1, 1, 2], [1, 2, 1], and [2, 1, 1] as distinct resamples. This method can provide a richer set of resamples, potentially enhancing the accuracy of the bootstrap estimate.

Furthermore, `exactamente` provides a standard bootstrap function where the user can specify the desired number of resamples, allowing for direct comparison between the exact methods and conventional bootstrap techniques.

## Installation

You can install the development version of `exactamente` from GitHub like so:

``` {r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("mightymetrika/exactamente")
```


## Example Usage

After installation, you can load the `exactamente` package using the library function.

```{r}
library(exactamente)
```


To utilize the fundamental tools provided by `exactamente`, begin by creating a numeric vector of data. You can then use the _bootstrap functions to obtain objects holding summary statistics and density estimates of the respective bootstrap distributions. In the following example, we use the [1, 2, 7] sample vector and take the mean as our test statistic.

```{r}
data <- c(1, 2, 7)

# Run exact case bootstrap
ec_res <- ecase_bootstrap(data)

# Run exact bootstrap
e_res <- exact_bootstrap(data)

# Run regular bootstrap
set.seed(183)
r_res <- reg_bootstrap(data)
```


Each _bootstrap function comes with plot and summary methods for further investigation of the bootstrap results.

```{r}
res <- list(ec_res, e_res, r_res)

lapply(res, plot)
lapply(res, summary)
```


`exactamente` also includes the `e_vs_r()` function, which enables direct comparison between the three methods using the same data set.

```{r}
set.seed(183)
comp_res <- e_vs_r(data)
comp_res$comp_plot
comp_res$summary_table
```


## Interactive Shiny App

One of the exciting features of the `exactamente` package is the inclusion of an interactive Shiny app. The app allows you to visually explore and compare different bootstrap methods. It is designed with a user-friendly interface and offers real-time results visualization.

You can access this app with the following command:

```{r, eval=FALSE}
exactamente_app()
```


## References

Jekel, C. F., & Romero, V. J. (2020). Conservative Estimation of Tail Probabilities from Limited Sample Data.
